PHQ4R - Data Cleaning (Study 2)

### 2024-01-30 study2/analysis/1_cleaning.qmd
update cleaning

---
title: "PHQ4R - Data Cleaning"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
---

@@

```{r}
df <- read.csv("../data/data_raw.csv")
```

@@

```{r}
# Consecutive count of participants per day (as area)
df |>
  mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
  arrange(Date) |>
  mutate(N = 1:nrow(df)) |>
  ggplot(aes(x = Date, y = N)) +
  geom_area() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    title = "Recruitment History",
    x = "Date",
    y = "Total Number of Participants"
  ) +
  see::theme_modern()
```

@@

```{r}
recode_phq <- function(x) {
  case_when(
    x == 1 ~ 0.5,
    x == 2 ~ 1,
    x == 3 ~ 2,
    x == 4 ~ 3,
    .default = x
  )
}

@@

```{r}
df |>
  select(matches("PHQ4_")) |>
  pivot_longer(cols = matches("Depression_|Anxiety_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count") +
  facet_grid(item ~ PHQ4_Condition) +
  labs(x = "Score", y = "N answers")
```

@@

```{r}
df |>
  select(matches("STAI5_"), -STAI5_Duration, -STAI5_Order) |>
  pivot_longer(cols = matches("STAI5_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count") +
  facet_wrap(~item) +
  labs(x = "Score", y = "N answers")
```

```{r}
df <- df |> 
  mutate(STAI5_General = (STAI5_1 + STAI5_2 + STAI5_3 + STAI5_4 + STAI5_5) / 5)
```

@@

```{r}
df |>
  select(matches("BDI2_"), -BDI2_Duration, -BDI2_Order) |>
  pivot_longer(cols = matches("BDI2_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count") +
  facet_wrap(~item) +
  labs(x = "Score", y = "N answers")
```

```{r}
df <- df |> 
  mutate(BDI2_Total = rowSums(select(df, matches("BDI2_"), -BDI2_Duration, -BDI2_Order)))
```

## Outliers

@@

```{r}
df |>
  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 50, "extra", "ok"),
         Duration = ifelse(Experiment_Duration > 50, 50, Experiment_Duration)) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_point(aes(color = Category, shape = Category)) +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  scale_color_manual(values = c("extra" = "red", ok = "black")) +
  guides(color = "none", shape = "none") +
  ggside::geom_xsidedensity(fill = "grey", color=NA) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Experiment Completion Time",
    x = "Duration (in minutes)",
    y = "Participant"
  )  +
  see::theme_modern() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3) 
```

## Descriptive Statistics

@@

```{r}
df |> 
  select(PHQ4_Condition, 
         STAI5 = STAI5_General,
         BDI2 = BDI2_Total,
         PHQ4_Total) |> 
  standardize() |> 
  estimate_density(at = "PHQ4_Condition") |> 
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color = Parameter)) +
  labs(x = "Standardized Score", y = "Distribution") +
  facet_wrap(~PHQ4_Condition)
```

>>>>

---
title: "PHQ4R - Data Cleaning (Study 2)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
---

@@

df <- read.csv("https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data_raw.csv")
```

@@

```{r}
# Consecutive count of participants per day (as area)
# df |>
#   mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
#   arrange(Date) |>
#   mutate(N = 1:nrow(df)) |>
#   ggplot(aes(x = Date, y = N)) +
#   geom_area() +
#   scale_y_continuous(expand = c(0, 0)) +
#   labs(
#     title = "Recruitment History",
#     x = "Date",
#     y = "Total Number of Participants"
#   ) +
#   see::theme_modern()
```

```{r}
#| message: false

# Consecutive count of participants per day (as area)
df |>
  mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
  group_by(Date, Experimenter) |> 
  summarize(N = n()) |> 
  ungroup() |>
  complete(Date, Experimenter, fill = list(N = 0)) |> 
  group_by(Experimenter) |>
  mutate(N = cumsum(N)) |>
  ggplot(aes(x = Date, y = N)) +
  geom_area(aes(fill=Experimenter)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    title = "Recruitment History",
    x = "Date",
    y = "Total Number of Participants"
  ) +
  see::theme_modern()
```

@@

```{r}
#| echo: true
#| code-fold: show

recode_phq <- function(x) {
  case_when(
    x == 1 ~ 0.5,
    x == 2 ~ 1,
    x == 3 ~ 2,
    x == 4 ~ 3,
    .default = x
  )
}

@@

```{r}
df |>
  select(matches("PHQ4_")) |>
  pivot_longer(cols = matches("Depression_|Anxiety_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count", fill="#9C27B0") +
  facet_grid(item ~ PHQ4_Condition) +
  labs(x = "Score", y = "N answers") +
  theme_bw()
```

@@

```{r}
df |>
  select(matches("STAI5_"), -STAI5_Duration, -STAI5_Order) |>
  pivot_longer(cols = matches("STAI5_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count", fill="#E91E63") +
  facet_wrap(~item) +
  labs(x = "Score", y = "N answers") +
  theme_bw()
```

```{r}
#| echo: true
#| code-fold: show

df <- df |> 
  mutate(STAI5_General = (STAI5_1 + STAI5_2 + STAI5_3 + STAI5_4 + STAI5_5) / 5)
```

@@

```{r}
df |>
  select(matches("BDI2_"), -BDI2_Duration, -BDI2_Order) |>
  pivot_longer(cols = matches("BDI2_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count", fill="#2196F3") +
  facet_wrap(~item) +
  labs(x = "Score", y = "N answers") +
  theme_bw()
```

```{r}
#| echo: true
#| code-fold: show

df <- df |> 
  mutate(BDI2_Total = rowSums(select(df, matches("BDI2_"), -BDI2_Duration, -BDI2_Order)))
```

## Exclusions

@@

```{r}
df |>
  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 30, "extra", "ok"),
         Duration = ifelse(Experiment_Duration > 30, 30, Experiment_Duration)) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
  geom_point(aes(color = Category, shape = Category)) +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  scale_color_manual(values = c("extra" = "red", ok = "black")) +
  guides(color = "none", shape = "none") +
  ggside::geom_xsidedensity(fill = "grey", color=NA) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Experiment Completion Time",
    x = "Duration (in minutes)",
    y = "Participant"
  )  +
  theme_bw() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3,
        panel.border = element_blank()) 
```

## Final Sample

@@

```{r}
df |> 
  select(PHQ4_Condition, 
         STAI5 = STAI5_General,
         BDI2 = BDI2_Total,
         PHQ4_Total) |> 
  standardize() |> 
  estimate_density(at = "PHQ4_Condition") |> 
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color = Parameter), linewidth=2) +
  scale_color_manual(values = c("STAI5" = "#E91E63", "BDI2" = "#2196F3", "PHQ4_Total" = "#9C27B0")) +
  labs(x = "Standardized Score", y = "Distribution") +
  facet_wrap(~PHQ4_Condition) +
  theme_bw()
```
